---
title: 'STAD80: Assignment 2'
author: "Vladislav Trukhin"
date: 'Due: Feb 3'
output:
  pdf_document:
    keep_tex: true
    toc: yes
    toc_depth: 2
---
## Question 1
### 1a
```{r}
data <- read.csv("/Users/vladislavtrukhin/Downloads/_data_hw2/housingprice.csv")

mean_prices <- tapply(data$price, data$zipcode, mean)
sorted_mean_prices <- sort(mean_prices, decreasing = TRUE)
labels(sorted_mean_prices) # Zipcode order by avg housing price
top_mean_prices <- sorted_mean_prices[0:3]
labels(top_mean_prices) # Top 3 zipcodes with most expensive avg housing price

par(mfrow=c(1,3))
boxplot(data[which(data$zipcode == labels(top_mean_prices[1])),]$price)
title(labels(top_mean_prices[1]))
boxplot(data[which(data$zipcode == labels(top_mean_prices[2])),]$price)
title(labels(top_mean_prices[2]))
boxplot(data[which(data$zipcode == labels(top_mean_prices[3])),]$price)
title(labels(top_mean_prices[3]))
```

### 1b
```{r}
plot(data$sqft_living, data$price)
```

### 1c
```{r}
train <- read.csv("/Users/vladislavtrukhin/Downloads/_data_hw2/train.data.csv")
test <- read.csv("/Users/vladislavtrukhin/Downloads/_data_hw2/test.data.csv")

fit <- lm(price ~ bedrooms + bathrooms + sqft_living + sqft_lot, train)
summary(fit)$r.squared # Training R2
cor(predict(fit, test), test$price)^2 # Test R2
```

### 1d
```{r}
fit <- lm(price ~ zipcode + 
            bedrooms + 
            bathrooms + 
            sqft_living + 
            sqft_lot, train)
summary(fit)$r.squared # Training R2
cor(predict(fit, test), test$price)^2 # Test R2
```

### 1e
```{r}
fancy = read.csv("/Users/vladislavtrukhin/Downloads/_data_hw2/fancyhouse.csv")
predict(fit, fancy) # Predicted price
```

The predicted price is not reasonable as the actual price of the home is $100+ 
million, which makes the predicted 10 times off of the actual price.

### 1f
$R^2 = 1-\frac{RSS}{TSS}$

As the value of TSS is the same for both models, only need to observe their respective RSS.

$RSS_{d+1} - RSS_d = ||\textbf{Y} - \textbf{X}_{d+1}\hat{\beta_{d+1}}||^2_2 - ||\textbf{Y} - \textbf{X}_d\hat{\beta}_d||^2_2$

$= \sum^n_{i=1}(y_i-\hat{\beta}_{d+1,0}+\hat{\beta}_{d+1, 1}x_{i,1}+...+\hat{\beta}_{d+1, d+1}x_{i,d+1})^2 - \sum^n_{i=1}(y_i-\hat{\beta}_{d,0}+\hat{\beta}_{d, 1}x_{i,1}+...+\hat{\beta}_{d, d}x_{i,d})^2$

$= \sum^n_{i=1}(y_i-\hat{y_i}_{d+1})^2 - \sum^n_{i=1}(y_i-\hat{y_i}_d)^2$

The addition of one covariate term brings one more degree of freedom when 
finding the minimum argument $\hat{\beta}_{d+1}$ for $||\textbf{Y} - \textbf{X}_{d+1}\beta_{d+1}||^2_2$. 
The minimum argument can be computed in closed form without issues as it is assumed
$n > d+1$, or number of features + 1 do not exceed number of samples within the 
training data, $\textbf{X}_{d+1}$. This means that the estimated $\textbf{Y}$, $\hat{\textbf{Y}}_{d+1} = \textbf{X}_{d+1}\beta_{d+1}$, will be closer to the true value of $\textbf{Y}$ than if were using the minimum argument obtained from the model without the additional covariate term, $\hat{\textbf{Y}}_{d} = \textbf{X}_{d}\beta_{d}$.

$\Rightarrow (y_i-\hat{\beta}_{d+1,0}+\hat{\beta}_{d+1, 1}x_{i,1}+...+\hat{\beta}_{d+1, d+1}x_{i,d+1})^2 \leq (y_i-\hat{\beta}_{d,0}+\hat{\beta}_{d, 1}x_{i,1}+...+\hat{\beta}_{d, d}x_{i,d})^2$

$\Rightarrow \sum^n_{i=1}(y_i-\hat{\beta}_{d+1,0}+\hat{\beta}_{d+1, 1}x_{i,1}+...+\hat{\beta}_{d+1, d+1}x_{i,d+1})^2 - \sum^n_{i=1}(y_i-\hat{\beta}_{d,0}+\hat{\beta}_{d, 1}x_{i,1}+...+\hat{\beta}_{d, d}x_{i,d})^2 \leq 0$

$\Rightarrow ||\textbf{Y} - \textbf{X}_{d+1}\hat{\beta_{d+1}}||^2_2 - ||\textbf{Y} - \textbf{X}_d\hat{\beta}_d||^2_2 \leq 0$

$\Rightarrow RSS_{d+1} - RSS_d \leq 0$

$\Rightarrow RSS_{d+1} \leq RSS_d$

$\Rightarrow \frac{RSS_{d+1}}{TSS} \leq \frac{RSS_d}{TSS}$

$\Rightarrow 1-\frac{RSS_{d+1}}{TSS} \geq 1-\frac{RSS_d}{TSS}$

$\Rightarrow R^2_{d+1} \geq R^2_d$

Therefore, if $n > d+1$, adding an additional covariate never lowers $R^2$ over 
training data.

## Question 2
### 2a 
```{r}
fit <- lm(price ~ zipcode + 
            bedrooms + 
            bathrooms + 
            bedrooms * bathrooms + 
            sqft_living + 
            sqft_lot, train)
summary(fit)$r.squared # Training R2
cor(predict(fit, test), test$price)^2 # Test R2
```

### 2b
```{r}
fit <- lm(price ~ zipcode + 
            bedrooms + 
            bathrooms + 
            bedrooms * bathrooms + 
            sqft_living + 
            sqft_lot + 
            sqft_living * bedrooms, train)
summary(fit)$r.squared # Training R2
cor(predict(fit, test), test$price)^2 # Test R2
```

Adding a new covariate that multiplies sqft_living and bedrooms. It models both 
the number of bedrooms and the size of each, which may influence the price.

### 2c
```{r}
fit <- lm(price ~ zipcode + 
            poly(bedrooms, 2) + 
            poly(bathrooms, 3) + 
            sqft_living + 
            sqft_lot, train)
summary(fit)$r.squared # Training R2
cor(predict(fit, test), test$price)^2 # Test R2
```

## Question 3
### 3.a
```{r}
wine_train <- 
  read.csv("/Users/vladislavtrukhin/Downloads/_data_hw2/wine.csv")
wine_test <- 
  read.csv("/Users/vladislavtrukhin/Downloads/_data_hw2/winetest.csv")
par(mfrow=c(2,2))
plot(wine_train$AGST, wine_train$Price)
plot(wine_train$WinterRain, wine_train$Price)
plot(wine_train$HarvestRain, wine_train$Price)
plot(wine_train$Age, wine_train$Price)

cor(wine_train$AGST, wine_train$Price)
cor(wine_train$WinterRain, wine_train$Price)
cor(wine_train$HarvestRain, wine_train$Price)
cor(wine_train$Age, wine_train$Price)
```

According to the graph, AGST and Price seem to be the most correlated. The 
variance is smaller and has a strong positive correlation. The Pearson's 
correlation number suggests the same, with the magnitude of the value for AGST 
and Price higher than all three.

### 3.b
```{r}
fit <- lm(Price ~ AGST, wine_train)
summary(fit)$coeff # Coefficients
summary(fit)$r.squared # Training R2
rss <- sum((predict(fit,wine_test)-wine_test$Price)^2)
tss <- sum((wine_test$Price-mean(wine_test$Price))^2)
rsq <- 1 - rss/tss
rsq # Test R2
```

### 3.c
```{r}
fit <- lm(Price ~ AGST + HarvestRain, wine_train)
summary(fit)$r.squared # Training R2
rss <- sum((predict(fit,wine_test)-wine_test$Price)^2)
tss <- sum((wine_test$Price-mean(wine_test$Price))^2)
rsq <- 1 - rss/tss
rsq # Test R2

fit <- lm(Price ~ AGST + HarvestRain + Age, wine_train)
summary(fit)$r.squared # Training R2
rss <- sum((predict(fit,wine_test)-wine_test$Price)^2)
tss <- sum((wine_test$Price-mean(wine_test$Price))^2)
rsq <- 1 - rss/tss
rsq # Test R2

fit <- lm(Price ~ AGST + HarvestRain + Age + WinterRain, wine_train)
summary(fit)$r.squared # Training R2
summary(fit)$coeff # Coefficients 
rss <- sum((predict(fit,wine_test)-wine_test$Price)^2)
tss <- sum((wine_test$Price-mean(wine_test$Price))^2)
rsq <- 1 - rss/tss
rsq # Test R2

fit <- lm(Price ~ AGST + HarvestRain + Age + WinterRain + FrancePop, wine_train)
summary(fit)$r.squared # Training R2
rss <- sum((predict(fit,wine_test)-wine_test$Price)^2)
tss <- sum((wine_test$Price-mean(wine_test$Price))^2)
rsq <- 1 - rss/tss
rsq # Test R2
```

The linear model depending on AGST, HarvestRain, Age, and WinterRain performed 
the best, as it had a high $R^2$ value for the training data and the highest 
$R^2$ value for the test data. That particular model agrees with 
Prof. Ashenfelter’s findings, since HarvestRain has a negative coefficient
and WinterRain has a positive one.

## Question 4
### 4.a
```{r}
baseball = read.csv("/Users/vladislavtrukhin/Downloads/_data_hw2/baseball.csv")
par(mfrow=c(1, 2))
boxplot(baseball$OBP)
hist(baseball$OBP)
mean(baseball$OBP)
median(baseball$OBP)

boxplot(baseball$SLG)
hist(baseball$SLG)
mean(baseball$SLG) 
median(baseball$SLG) 

boxplot(baseball$BA)
hist(baseball$BA)
mean(baseball$BA)
median(baseball$BA)
```

### 4.b
```{r}
par(mfrow=c(1,2))
fit <- lm(baseball$RS ~ baseball$BA)
summary(fit)$coeff # Coefficients
summary(fit)$r.squared # Training R2
plot(baseball$BA, baseball$RS)
abline(fit)
qqnorm(rstandard(fit))

fit <- lm(baseball$RS ~ baseball$OBP)
summary(fit)$coeff # Coefficients
summary(fit)$r.squared # Training R2
plot(baseball$OBP, baseball$RS)
abline(fit)
qqnorm(rstandard(fit))

fit <- lm(baseball$RS ~ baseball$SLG)
summary(fit)$coeff # Coefficients
summary(fit)$r.squared # Training R2
plot(baseball$SLG, baseball$RS)
abline(fit)
qqnorm(rstandard(fit))
```

The analysis is not consistent with the intuition, as the $R^2$ is the 
lowest relative to RS's and BA's $R^2$.

### 4.c
```{r}
fit <- lm(baseball$RS ~ baseball$BA + baseball$SLG + baseball$OBP)
summary(fit)$coeff # Coefficients
summary(fit)$r.squared # Training R2
qqnorm(rstandard(fit))

fit <- lm(baseball$RS ~ baseball$OBP + baseball$SLG)
summary(fit)$r.squared # Training R2
```

The results are consistent of that in 4.b. The coefficient of BA has a low 
significance level, consistent with the low $R^2$ value obtained from 4.b.
The two models have near equivalent $R^2$ values, which makes the later model 
a better model due to being more simple.

### 4.d
```{r}
baseball$RD = baseball$RS - baseball$RA
baseball_pre_2002 = baseball[which(baseball$Year < 2002), ]

fit1 <- lm(W ~ RD, baseball_pre_2002)
fit2 <- lm(RS ~ OBP + SLG, baseball_pre_2002)
fit3 <- lm(RA ~ OOBP + OSLG, baseball_pre_2002)

oak_pred <- data.frame("OBP" = .349, 
                       "SLG" = .430, 
                       "OOBP" = .307, 
                       "OSLG" = .373)
oak_pred$RS <- predict(fit2, oak_pred)
oak_pred$RA <- predict(fit3, oak_pred)
oak_pred$RD <- oak_pred$RS - oak_pred$RA
predict(fit1, oak_pred) #Predicted Wins
baseball[which(baseball$Year == 2002 & baseball$Team == 'OAK'), ]$W #Actual Wins
```